---
title: "What are we doing?"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(DT)
opts_chunk$set(echo = TRUE,
               message = FALSE,
               warning = FALSE)
```

## Categories

```{r catTable, echo=FALSE}
df <- data.frame(
  category = c("Bacteria Contamination","Bacteria Type",
               "Geography","Season", 
               "Flow Type I", "Flow Type II", 
               "Flow Conditions"),
  choices = c("1-6","Human, Uncontaminated, Dog, Animal, Atypical human, Uncertain",
              "WI, MI, NY","Fall, Winter, Spring, Summer",
              "Environmental, Sewage","Stream, Storm Sewer, Sewage", "Base, Event"),
  stringsAsFactors = FALSE
)

kable(df)

```

Most of those categories are self-explanatory, but categorizing the bacteria contamination was not:

### Bacteria Contamination

This was tricky because there were 3 measured bacteria values, two measurements of lachno (Lachnoclostridium?), and one of bacHum (?). For the contamination categories, we ignored lachno2 (this was used though in the "Bacteria Type" categories). 

First, we classified each of those bacteria levels. Then, we combined the 2 classifications. Hopefully this chart explains that:

First, both lachno and bachHum were classified:

```{r firstClass, echo=FALSE}
df_hum <- data.frame(
  measurement = c("0-225","226-1,000","1,001-10,000", "10,001-100,000", "100,001-1,000,000", ">1,000,000"),
  classification = c("ND","Low","Medium","High","Very High", "Extreme"),
  stringsAsFactors = FALSE
)

kable(df_hum)

```

Now to combine:
```{r comboClass, echo=FALSE}
# Here's a table...but I think the graph is much easier to follow:
# df_combo <- data.frame(
#   category = c("1","2","","3","","","4","","","","5","","6"),
#   lower_bound = c("ND",
#                   "ND",
#                   "low",
#                   "ND",
#                   "low",
#                   "medium",
#                   "ND",
#                   "low",
#                   "medium",
#                   "high",
#                   "low to medium",
#                   "high to very-high",
#                   "high to extreme"),
#   upper_bound = c("ND",
#                   "low to medium",
#                   "low",
#                   "high",
#                   "medium to high",
#                   "medium",
#                   "very-high to extreme",
#                   "very-high",
#                   "high to very-high",
#                   "high",
#                   "extreme",
#                   "very-high",
#                   "extreme"),
#   stringsAsFactors = FALSE
# )
# kable(df_combo)
```

```{r tilesCat, echo=FALSE}
library(ggplot2)

df_2 <- expand.grid(c("ND","Low","Medium","High","Very High", "Extreme"),
                    c("ND","Low","Medium","High","Very High", "Extreme"))
df_2$Category <- NA
df_2$Category <- c("1","2","2","3","4","4","2","2",
              "3","3","4","5","2","3","3","4",
              "4","5","3","3","4","4","5","6",
              "4","4","4","5","5","6","4","5",
              "5","6","6","6")

ggplot(data = df_2) +
  geom_tile(aes(x = Var1, y=Var2, fill = Category)) +
  theme_minimal() +
  theme(axis.title = element_blank())

```

If the measurements were very close to a cutoffs, we'd explore nudging them up or down to see how the results were affected.

### Bacteria Type

Additionally, for some of the data, we had a "dog" marker. This allowed a further characterization from the bacteria measurements. If this is needed to be described, it might be easiest with a tree diagram.

## Optical markers

Next step was to create the optical categories. 

### Fluorescence pairs and means

```{r flPairs, echo=FALSE}
DT::datatable(read.csv("raw_data/opticalSummary/ex_ems_meansCA.csv",as.is=TRUE), rownames = FALSE, options = list(pageLength = 5))
```

### HIX, FI, Freshness

"Computes humification index and fluorescence indes from fluorescence data. HIX as defined by Ohno, 2002, Fluorescence inner-filtering correction for determining the humification index of dissolved organic matter"

### Single absorbance signals

The following wavelengths are flagged from the absorbance measurements: `r paste(read.csv("raw_data/opticalSummary/abs_wavsCA.csv",as.is=TRUE)[,1], collapse = ", ")`.

### Spectral slopes

```{r spectralSlopes, echo=FALSE}
DT::datatable(read.csv("raw_data/opticalSummary/SagWavesCA.csv",as.is=TRUE), rownames = FALSE)
```

### Deviance of absorbance from exponential regression in the wastewater area

"Computes residuals from a linear regression using the first order decay function as defined in Helms et al. 2008, Limnol. Oceanogr., 53(3), 955-969"

```{r devience, echo=TRUE, eval=FALSE}
getExpResid(wavelength=269,
           rangeReg=c(240,341),rangeGap=c(254,302),
           dataAbs=dfabs,
           waveCol="nm",colSubsetString="gr",
           dataSummary=dfOpt,grnum="CAGRnumber")
```

### Ratios

```{r ratios, echo=FALSE}
orderRatios <- function(sigs){
    #Ratios of a few things
    ratioVars <- expand.grid(sigs,sigs, stringsAsFactors = FALSE) %>%
      filter(Var1 != Var2) 
    
    ratioVars$lookup <- apply(ratioVars[,c("Var1","Var2")],1, function(x){paste(sort(x),collapse="~")})
    
    ratioVars <- filter(ratioVars, !duplicated(lookup)) %>%
      select(-lookup)    
    return(ratioVars)
  }

SummaryDir <- "raw_data/opticalSummary/"
ratioSignalsAbs <- read.csv(file.path(SummaryDir,"ratioSignalsAbsCA.csv"),as.is=TRUE)
ratioSignalsAbs <- ratioSignalsAbs[which(ratioSignalsAbs[2]>0),1]
ratioSignalsSr <- read.csv(file.path(SummaryDir,"ratioSignalsSrCA.csv"),as.is=TRUE)
ratioSignalsSr <- ratioSignalsSr[which(ratioSignalsSr[2]>0),1]
ratioSignalsSniff <- read.csv(file.path(SummaryDir,"ratioSignalsSniff.csv"),as.is=TRUE)
ratioSignalsSniff <- ratioSignalsSniff[which(ratioSignalsSniff[2]>0),1]
ratioSignalsSniffWetStar <- read.csv(file.path(SummaryDir,"ratioSignalsSniffAll.csv"),as.is=TRUE)[,1]
logSignals <- read.csv(file.path(SummaryDir,"logSignalsCA.csv"),as.is=TRUE)

ratios <- dplyr::bind_rows(orderRatios(ratioSignalsAbs), 
  orderRatios(ratioSignalsSr), 
  orderRatios(ratioSignalsSniff), 
  orderRatios(c("Sn.1","Sn.2","Sn.3","Sn.4",                                               "Sn.5","Sn.6","Sn.7","Sn.8","Sn.9")))

DT::datatable(ratios, rownames = FALSE, options = list(pageLength = 5)) %>%
  formatStyle(colnames(ratios), `text-align` = 'center')

```

### Ratios minus baseline

```{r ratiosWBaseline, echo=FALSE}

df_1 <- orderRatios(ratioSignalsAbs[ratioSignalsAbs != "A254"])
df_1$baseline <- "A254"
df_2 <- orderRatios(ratioSignalsSniff[ratioSignalsSniff != "F"])
df_2$baseline <- "F"

ratio_base <- dplyr::bind_rows(df_1, df_2)

DT::datatable(ratio_base, rownames = FALSE, options = list(pageLength = 5)) %>%
  formatStyle(colnames(ratio_base), `text-align` = 'center')

```

### Log Transform

`r paste(logSignals[,1], collapse = ", ")`

## Field measurements

Some sites had pH, DO, turbidity, specific conductance, and water temperature, measurements as well. When possible we tried to incorporate these in models.
